---
title: "Automatically Get Data From ATL PD Data Downloads Page"
output: html_notebook
---

### Description
The Atlanta Police Department (APD) publishes data on a weekly basis at http://www.atlantapd.org/i-want-to/crime-data-downloads.  Files are separated by year, and the more recent years' datasets are .zip archives; older years are published as .xlsx files.  This code automatically downloads the data and combines it into a single CSV file.

```{r, message=FALSE}
library(dplyr); library(rvest); library(stringr); library(lubridate);
library(readr); library(httr); library(readxl)
```

### Get metadata of the download links

```{r Get Crime Link Metadata}
crime_links_html <- 
     read_html("http://www.atlantapd.org/i-want-to/crime-data-downloads") %>% 
     html_nodes("#widget_362_0_197 .content_link")

crime_links <- 
     data_frame(filename = crime_links_html %>% html_text(),
           url = crime_links_html %>% html_attr("href") %>% paste0("http://www.atlantapd.org", .))

temp_list <- vector(mode = "list", length = 9)
for(i in seq_along(crime_links$url)){
     temp_list[[i]] <- headers(GET(crime_links$url[i]))     
}

crime_links_fileinfo <- 
     lapply(temp_list, function(x) data_frame(type = x$`content-type`,
                                         disposition = x$`content-disposition`, 
                                         documenttitle = x$`content-documenttitle`,
                                         lastmodified = x$`last-modified`)) %>%
     bind_rows()

crime_links <- bind_cols(crime_links, crime_links_fileinfo) %>% 
     mutate(filename2 = str_sub(disposition, start = 11, end = -2))

crime_links
```

### Functions for downloading data files

```{r Data download functions}
get_crime_zipfile <- function(zip.url){
     td <- tempdir()
     print(paste0("Reading ", zip.url))
     tf <- tempfile(tmpdir=td, fileext=".zip") 
     download.file(zip.url, tf, mode = "wb")
     fname <- unzip(tf, list=TRUE)$Name[1] 
     unzip(tf, files=fname, exdir=td,overwrite=TRUE)
     fpath <- file.path(td, fname)
     print(excel_sheets(fpath))
     read_excel(fpath, sheet = "Query")
}

get_crime_nonzipfile <- function(file.url, file.name){
     print(paste0("Downloading file: ", file.name, ", URL: ", file.url))
     download.file(url = file.url, destfile = file.name, mode = "wb")
     print(excel_sheets(file.name))
     return(read_excel(file.name, sheet="Query"))
}

```

### Download data & combine into one dataset

```{r}
crime_data <- vector(mode = "list", length = nrow(crime_links))
for(i in seq_along(crime_links$filename)){
     if(crime_links$type[i] == "application/zip"){
          crime_data[[i]] <- get_crime_zipfile(crime_links$url[i])
     }else if(crime_links$type[i] == "application/vnd.openxmlformats-officedocument.spreadsheetml.sheet"){
          crime_data[[i]] <- 
          get_crime_nonzipfile(crime_links$url[i], file.name = crime_links$filename2[i])
     }
}

xlsx_filenames <- crime_links %>% .[str_detect(.$filename2, "\\.xlsx"),] %>% .$filename2
for(i in seq_along(xlsx_filenames)){
     unlink(xlsx_filenames[i])
}

crime_data2 <- 
     bind_rows(crime_data) %>%
     mutate(rpt_date = mdy(rpt_date), 
            occur_datetime = mdy_hms(paste(occur_date, occur_time)), occur_date = mdy(occur_date),
            poss_datetime = mdy_hms(paste(poss_date, poss_time)), poss_date = mdy(poss_date),
            MaxOfnum_victims = as.integer(MaxOfnum_victims),
            x = round(as.numeric(x), 5), y = round(as.numeric(y), 5))
```

### Write crime data & link metadata to file
- Write the crime data to `Data/crime_data.csv` & `Data/crime_data.rds` (RDS is a file type used in R, that preserves data types of the elements in a data frame, making reading & writing easier)
- Write link metadata to `Data/crime_links.csv`
```{r Write data to files}
write_rds(crime_data2, "Data/crime_data.rds")
write_csv(crime_data2, "Data/crime_data.csv")
write_csv(crime_links, "Data/crime_links.csv")
```